A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2018-10-23 doc/large_strain.tex
few fixes based on JPs comments

  w_q = \widetilde{w}_{q_1} \widetilde{w}_{q_2}

@@

Since each of these $2\times d$ loops has a complexity of $n$ and we perform them for $n^d$ values in quadrature points or values in degrees of freedoms simultaneously, we end up with an arithmetic complexity of $\mathcal{O}(dn^{d+1})$.
Hence, we can expect that such a matrix-free approach is faster than a regular matrix-based approach, especially for 3D and high polynomial degree with the corresponding quadrature rule.

@@

\begin{algorithm}[!ht]
  \SetKwInOut{Input}{Given}
  \SetKwInOut{Output}{Return}
  \Input{RHS FE vector $\gz x$,
  cached
  $c_1 := 2\left[\mu - 2 \lambda \log(J)\right]/J$,
  $c_2 := 2\lambda/J$ and $\gz \tau/J$ for each cell and quadrature point,
  evaluated based on $\overline{\gz u^h}$}
  \Output{action of the FE tangent \eqref{eq:algebraic_tangent} on $\gz x$}
  \ForEach{ element $K \in \Omega^h$ }{
        evaluate the following tensors at each quadrature point using sum factorization \;
        $\gz g_s := \gz \nabla_{sym} \gz x$ \tcp*{2nd order symmetric}
        $\gz g := \gz \nabla \gz x$ \tcp*{2nd order}
        \ForEach{quadrature point $q$ on $K$}{
          evaluate $\boldsymbol{\mathcal{G}}\gz g := \gz g \cdot \left[\gz \tau/J \right]$ \tcp*{2nd order}
          queue $\boldsymbol{\mathcal{G}} \gz g$ for contraction $\gz \nabla \gz N_i : \boldsymbol{\mathcal{G}} \gz g$ \;
          evaluate $\boldsymbol{\mathcal{C}} \gz g_s := c_1 \gz g_s + c_2 \, \rm{tr}(\gz g_s) \gz I$ \tcp*{2nd order symmetric}
          queue $\boldsymbol{\mathcal{C}} \gz g_s$ contraction $\gz \nabla_{sym} \gz N_i : \boldsymbol{\mathcal{C}} \gz g_s $ \;
        }
        evaluate queued contractions using sum factorization and distribute results to the destination vector
  }
  \caption{Matrix-free tangent cell operator, variant 1: cache second-order Kirchhoff stress $\gz \tau$ and thereby avoid the need to evaluate refential quantities like $\gz F$ at runtime.}
  \label{alg:mf_tensor2}
\end{algorithm}

>>>>

  w_q = \widetilde{w}_{q_1} \widetilde{w}_{q_2} \, .

@@

Since each of these $2\times d$ loops has a complexity of $n$ and we perform them for $n^d$ values in quadrature points or values in degrees of freedoms simultaneously, what results is an algorithm with arithmetic complexity of $\mathcal{O}(dn^{d+1})$.
Hence, we can expect that such a matrix-free approach is faster than a regular matrix-based approach, especially for the 3D case, and where the discretization and evaluation employs high polynomial degree basis functions with the corresponding quadrature rule.

@@

\begin{algorithm}[!ht]
  \SetKwInOut{Input}{Given}
  \SetKwInOut{Output}{Return}
  \Input{RHS FE vector $\gz x$,
  cached
  $c_1 := 2\left[\mu - 2 \lambda \log(J)\right]/J$,
  $c_2 := 2\lambda/J$ and $\gz \tau/J$ for each cell and quadrature point,
  evaluated based on $\overline{\gz u^h}$}
  \Output{action of the FE tangent \eqref{eq:algebraic_tangent} on $\gz x$}
  \ForEach{ element $K \in \Omega^h$ }{
        evaluate the following tensors at each quadrature point using sum factorization \;
        $\gz g_s := \gz \nabla_{sym} \gz x$ \tcp*{2nd order symmetric}
        $\gz g := \gz \nabla \gz x$ \tcp*{2nd order}
        \ForEach{quadrature point $q$ on $K$}{
          evaluate $\boldsymbol{\mathcal{G}}\gz g := \gz g \cdot \left[\gz \tau/J \right]$ \tcp*{2nd order}
          queue $\boldsymbol{\mathcal{G}} \gz g$ for contraction $\gz \nabla \gz N_i : \boldsymbol{\mathcal{G}} \gz g$ \;
          evaluate $\boldsymbol{\mathcal{C}} \gz g_s := c_1 \gz g_s + c_2 \, \rm{tr}(\gz g_s) \gz I$ \tcp*{2nd order symmetric}
          queue $\boldsymbol{\mathcal{C}} \gz g_s$ contraction $\gz \nabla_{sym} \gz N_i : \boldsymbol{\mathcal{C}} \gz g_s $ \;
        }
        evaluate queued contractions using sum factorization and distribute results to the destination vector
  }
  \caption{Matrix-free tangent cell operator, variant 1: cache second-order Kirchhoff stress $\gz \tau$ and thereby avoid the need to evaluate referential quantities like $\gz F$ at when executed.}
  \label{alg:mf_tensor2}
\end{algorithm}
